---
title: "Lab9_Covid19"
author: "Joe Drake"
date: "4/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)



time_series_confirmed_long2 <- read_csv(url("https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")

head(report_03_11_2020)

```
```{r}
str(report_03_11_2020)
```

#Note that older data from older reports are in different format than new format recently released:

```{r}
report_03_23_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-23-2020.csv"))
head(report_03_23_2020)
```


```{r}
str(report_03_23_2020)
```

#confirmed cases in US States as of March 11th

```{r}
report_03_11_2020 %>% 
  filter(Country_Region=="US") %>% 
  ggplot(aes(x=Confirmed, y=reorder(Province_State, Confirmed))) + geom_point() +
  ggtitle("Confirmed cases for each US State") +
  ylab("Country/Region") +
  xlab("Confirmed Cases")
```

```{r}
report_03_11_2020 %>% 
  group_by(Country_Region) %>% 
  summarise(Deaths=sum(Deaths)) %>% 
  arrange(desc(Deaths))
```

Countries with most cases/deaths

```{r}
report_03_11_2020 %>% 
  group_by(Country_Region) %>% 
  summarise(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:20) %>% 
  ggplot(aes(x=Deaths, y=reorder(Country_Region, Deaths))) +
    geom_bar(stat="identity") +
    ggtitle("The 20 countries with the most reported Cov19-related deaths") +
    ylab("Country/region") +
    xlab("Deaths")
  
```


##Timeseries
as of march 26, still useing old column headers
```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

head(time_series_confirmed)

```
Convert from wide to long data format
```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Confirmed")

head(time_series_confirmed_long)

```
# create another table of deaths and and join tables

```{r}
time_series_deaths <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths" )
head(time_series_confirmed_long)

```

We need a unique key column to successfully join the 2 tables
Use unite code to combine a couple columns to make a unique key column

```{r}
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep=".", remove=FALSE)
head(time_series_confirmed_long)

```
Can do the same for deaths while eliminating redudant columns

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep=".") %>% 
  select(Key, Deaths)

#use a full_join and use select to remove key

time_series_long_joined <- full_join(time_series_confirmed_long, time_series_deaths_long, by=c("Key"))

head(time_series_long_joined)
```

Check to make sure counts or deaths don't have NA

```{r}
which(is.na(time_series_long_joined$Confirmed))


which(is.na(time_series_long_joined$Deaths))
```

Add the number of recovered individuals

```{r}
### download the file
time_series_recovered <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

### convert to long format
time_series_recovered_long <- time_series_recovered %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

### Create the Key
time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)

### Merge with prior table (only this time will delete the Key column
### because we have no further use for it)
time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)
head(time_series_long_joined)
```

# NOW TO REFORMAT that ugly DATE

```{r}
time_series_long_joined$Date <-mdy(time_series_long_joined$Date)
```

We may want all the counts on the same graph, although there are multiple ways to accomplish this, one way is by creating a column 

```{r}
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),names_to = "Report_Type", values_to = "Counts")

head(time_series_long_joined_counts)
tail(time_series_long_joined_counts)

```

# Graphs from times series

US has to be summarize to get info from all states

```{r}
time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
  filter(Country_Region=="US") %>% 
  ggplot(aes(x=Date, y=Confirmed)) +
    geom_point() +
    geom_line() +
    ggtitle("Confirmed US Covid- 19 Cases")
```

Looking at US in context of other countries

```{r}

time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
  filter(Country_Region %in% c("China","Japan", "Korea, South", "Italy", "Spain", "US")) %>% 
  ggplot(aes(x=Date, y=Confirmed)) +
  geom_point() +
  geom_line() +
  ggtitle("Confirmed Covid-19 Cases") +
  facet_wrap(~Country_Region, ncol=2, scales="free_y")

```

```{r}
time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
  filter(Country_Region %in% c("China","Japan", "Korea, South", "Italy", "Spain", "US", "France")) %>% 
  ggplot(aes(x=Date, y=Confirmed, color=Country_Region)) +
  geom_point() +
  geom_line() +
  ggtitle("Confirmed Covid-19 Cases") 

```

use a log y-scale for visualizing relationship of counts confirmed vs. recovered, but personally I think this hides the real nature of the data

```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")

time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = (Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")

```

##MAPS
#by Anisa Dhana

```{r}
library(maps)
library(viridis)
world <- map_data("world")
#set cutoffs for case #
mybreaks <- c(1,20,100,1000,50000,200000)

# data from april 2
report_04_02_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv"))


ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_confirmed, aes(x=Long, y=Lat, size=`4/2/20`, color=`4/2/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000-199999", "200000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000-199999", "200000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )


```



#by Chris Sutherland

```{r}
time_series_confirmed_long2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province_State = "Province/State", Country_Region = "Country/Region") %>%
    pivot_longer(-c(Province_State, Country_Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))


head(time_series_confirmed_long2)

#plot a static representation

static <- ggplot(subset(time_series_confirmed_long2, Date %in% seq(min(Date), max(Date), 14)),
                 aes(x=Long, y=Lat, size=cumulative_cases/1000)) +
  borders("world", colour=NA, fill="grey90") +
  theme_bw() +
  geom_point(shape=21, color='purple', fill="purple", alpha=0.5) +
  labs(title="Covid 19 Spread", x="", y="", size="Cases (x1000)") +
  coord_fixed(ratio=1) +
  facet_wrap(.~Date,ncol=1)
  


static

```

## Exercises
1.
```{r}

time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region %in% c( "US", "Poland")) %>% 
    ggplot(aes(x = Date,  y = (Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases") +
    facet_wrap(~Country_Region)

```
2.
```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region %in% c( "US", "Serbia")) %>% 
    ggplot(aes(x = Date,  y = (Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases") +
    facet_wrap(~Country_Region, scales="free_y")

```

1.

```{r}
# data from april 2
report_04_02_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv"))

report_04_02_2020 %>% 
  group_by(Country_Region) %>% 
  summarise(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:20) %>% 
  ggplot(aes(x=Deaths, y=reorder(Country_Region, Deaths))) +
    geom_bar(stat="identity") +
    ggtitle("The 20 countries with the most reported Cov19-related deaths") +
    ylab("Country/region") +
    xlab("Deaths")


```

2.
```{r}
# data from april 2
report_04_02_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv"))

report_04_02_2020 %>% 
  group_by(Country_Region) %>% 
  summarise(Recovered = sum(Recovered)) %>% 
  arrange(desc(Recovered)) %>% 
  slice(1:20) %>% 
  ggplot(aes(x=Recovered, y=reorder(Country_Region, Recovered))) +
    geom_bar(stat="identity") +
    ggtitle("The 20 countries with the most reported Cov19-related Recovery") +
    ylab("Country/region") +
    xlab("Recovered")

```

1.
```{r}
library(maps)
library(viridis)
world <- map_data("world")
#set cutoffs for case #
mybreaks <- c(1,20,100,1000,50000,200000)

ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_confirmed, aes(x=Long, y=Lat, size=`4/2/20`, color=`4/2/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000-199999", "200000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000-199999", "200000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )


```
2.
```{r}

time_series_confirmed_long2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province_State = "Province/State", Country_Region = "Country/Region") %>%
    pivot_longer(-c(Province_State, Country_Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
  filter(Country_Region %in% "US") %>% 
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))


head(time_series_confirmed_long2)

#plot a static representation

static <- ggplot(subset(time_series_confirmed_long2, Date %in% seq(min(Date), max(Date), 14)),
                 aes(x=Long, y=Lat, size=cumulative_cases/1000)) +
  borders("usa", colour=NA, fill="grey90") +
  theme_bw() +
  geom_point(shape=21, color='purple', fill="purple", alpha=0.5) +
  labs(title="Covid 19 Spread", x="", y="", size="Cases (x1000)") +
  coord_fixed(ratio=1) +
  facet_wrap(.~Date,ncol=1)
  


static



```

